Simulations, CLT and EVT

### 2020-08-27 mean_null.Rmd
adding CLT, EVT and simulations

Samples are taken from Gamma(0.5, 3) distribution and $k = 500$ in all the cases. Small values of n considered in the range  ${2, 3, \ldots, 10}$ and large values of $n$ considered within ${25, 30, 35, \ldots, 45}$. We study the distribution of:

 1. sample means for small $n$

 2. sample means for large $n$

 3. normalised sample means for small $n$

 4. normalised sample means for large $n$

 5. sample maximums for small $n$

 6. sample maximums for large $n$

 7. normalised sample maximums for small $n$

 8. normalised sample maximums for large $n$

@@

```{r functions}
# create data for each levels
create_simdata <- function(nx = 2,
                           # number of rvs to generate
                           nsim = 500, 
                           # number of simulations
                           sim_dist = distributional::dist_gamma(0.5, 3),
                           # distribution of each rv
                           create_fun = mean 
                           # function used on the rvs 
                           ){
  mean_dist <- (1:nsim) %>%
    purrr::map(function(i) {
      x <- sim_dist %>%
        distributional::generate(nx) %>%
        unlist()
      x %>% create_fun()
    })
}
```

@@

plot_simdata <- function(nlevels = seq(2, 10, 1),
                         nsim = 500, 
                         sim_dist = distributional::dist_gamma(0.5, 3),
                         create_fun = mean, 
                         distr = "norm"){
data_orig <- nlevels %>%
  purrr::map_df(function(i) {
    create_datai <- create_simdata(nx = i, nsim, sim_dist, create_fun)
    tibble::tibble(ind = i, sim_data = create_datai)
  })

# defining gumbel distribution
dgumbel <- function(x, a, b) 1/b*exp((a-x)/b)*exp(-exp((a-x)/b))
pgumbel <- function(q, a, b) exp(-exp((a-q)/b))
qgumbel <- function(p, a, b) a-b*log(-log(p))

response <- data_orig %>%
  unnest(sim_data)

if(distr == "gumbel")
{
par = fitdistrplus::fitdist(response$sim_data, distr = "gumbel", start = list(a = 0, b = 1))
}

if(distr == "norm")
{
par = fitdistrplus::fitdist(response$sim_data, distr = distr)
}

par1 <- round(par$estimate[1], 2)
par2 <- round(par$estimate[2], 2)

data_orig %>%
  unnest(sim_data) %>%
  ggplot(aes(x = sim_data)) +
  geom_histogram() +
  facet_wrap(~ind) + 
  ggtitle(paste0(distr, " (",par1,", ",par2,")"))
}
```

@@

```{r}
set.seed(12345)
plot_simdata(nlevels = seq(2, 10, 1),
             nsim = 500, 
             create_fun = mean,
             distr = "norm")
```

@@

```{r}
set.seed(54321)
plot_simdata(nlevels = seq(20, 45, 5),
             nsim = 500,
             create_fun = mean)
```

```{r norm_mean}
norm_mean <- function(x = NULL){
    shape = 0.5
    rate = 3
    meanx = shape/rate
    sdx = shape/(rate^2)
    (mean(x) - meanx)/sdx
}
```

@@

```{r normmean_small_lev}
set.seed(12345)
plot_simdata(nlevels = seq(2, 10, 1),
             nsim = 500, 
             create_fun = norm_mean)
```

@@

```{r normmean_large_lev}
set.seed(54321)
plot_simdata(nlevels = seq(20, 45, 5),
             nsim = 500, 
             create_fun = norm_mean,
             distr = "norm")
```

@@

```{r max_small_lev}
set.seed(12345)
plot_simdata(nlevels = seq(2, 10, 1),
             nsim = 500, 
             create_fun = max)
```

@@

```{r max_large_lev}
set.seed(54321)
plot_simdata(nlevels = seq(20, 45, 5),
             nsim = 500, 
             create_fun = max)
```

```{r norm_max}
# normalised maximum of a vector
norm_max <- function(x = NULL){
      nx <- length(x)
      p = 1-(1/nx)
      shape = 0.5
      rate = 3
      meanx = shape/rate
      sdx = shape/(rate^2)
      b = stats::qnorm(p, meanx, sdx)
      a = 1/(nx*stats::dnorm(b, meanx, sdx))
      #k = stats::quantile(x, p, type = 8, na.rm = TRUE)
      (max(x) - a)/b
      #max(x)/k
      }
```

@@

```{r normmax_small_lev}
set.seed(12345)
plot_simdata(nlevels = seq(2, 10, 1),
             nsim = 500, 
             create_fun = norm_max,
             distr = "gumbel")
```

@@

```{r normmax_large_lev}
set.seed(54321)
plot_simdata(nlevels = seq(20, 45, 5),
             nsim = 500, 
             create_fun = norm_max, 
             distr = "gumbel")
```

>>>>

Samples are taken from Gamma(3, 0.5) distribution and $k = 500$ in all the cases. Small values of n considered in the range  ${2, 3, \ldots, 10}$ and large values of $n$ considered within ${25, 30, 35, \ldots, 45}$. We study the distribution of:

 1. sample means for small $n$ 
(Obs: distribution doesn't look similar or normal)

 2. sample means for large $n$
(Obs: distribution looks similar and normal like with different scales)

 3. normalised sample means for small $n$
(Obs: distribution doesn't look similar)

 4. normalised sample means for large $n$ 
(Obs: distribution looks similar in both shape and scale with mean 0)

 5. sample maximums for small $n$ 
(Obs: distribution doesn't look too different except for tails, high scale)

 6. sample maximums for large $n$ 
(Obs: distribution looks similar and positively skewed with very high scales)

 7. normalised sample maximums for small $n$ 
(Obs: distribution doesn't look similar, scaled down, maybe similar after facet label 5)

 8. normalised sample maximums for large $n$ 
(Obs: distribution looks similar and it has been scaled down much more than maximum without normalisation. Can we use this fact to explore maximums more?)

Q: So when we are talking about normalising for the levels when number of levels are low? - Not by using limit theorems. Transformation?

@@

```{r shape_rate}
shape1 = 3
rate1 = 0.5
```

```{r functions}
# create data for each levels

create_simdata <- function(nx = 2,
                           # number of rvs to generate
                           nsim = 500,
                           # number of simulations
                           sim_dist = distributional::dist_gamma(shape = shape1, 
                                                                 rate = rate1),
                           # distribution of each rv
                           create_fun = mean
                           # function used on the rvs
) {
  mean_dist <- (1:nsim) %>%
    purrr::map(function(i) {
      x <- sim_dist %>%
        distributional::generate(nx) %>%
        unlist()
      x %>% create_fun()
    })
}
```

@@

plot_simdata <- function(nlevels = seq(2, 10, 1),
                         nsim = 500,
                         sim_dist = distributional::dist_gamma(shape = shape1,
                      rate = rate1),
                         # distr = "norm",
                         create_fun = mean) {
  data_orig <- nlevels %>%
    purrr::map_df(function(i) {
      create_datai <- create_simdata(nx = i, nsim, sim_dist, create_fun)
      tibble::tibble(ind = i, sim_data = create_datai)
    })

  data_orig %>%
    unnest(sim_data) %>%
    ggplot(aes(x = sim_data)) +
    geom_histogram() +
    facet_wrap(~ind)
  # + ggtitle(paste0(distr, " (",par1,", ",par2,")"))
}
```

@@

```{r}
set.seed(12345)
plot_simdata(
  nlevels = seq(2, 10, 1),
  nsim = 500,
  create_fun = mean
)
```

@@

```{r}
set.seed(54321)
plot_simdata(
  nlevels = seq(20, 45, 5),
  nsim = 500,
  create_fun = mean
)
```

```{r norm_mean}
norm_mean <- function(x = NULL,
                      shape = shape1,
                      rate = rate1) {
  meanx <- shape / rate
  sdx <- shape / (rate^2)
  (mean(x) - meanx) / sdx
}
```

@@

```{r normmean_small_lev}
set.seed(12345)
plot_simdata(
  nlevels = seq(2, 10, 1),
  nsim = 500,
  create_fun = norm_mean
)
```

@@

```{r normmean_large_lev}
set.seed(54321)
plot_simdata(
  nlevels = seq(20, 45, 5),
  nsim = 500,
  create_fun = norm_mean
)
```

@@

```{r max_small_lev}
set.seed(12345)
plot_simdata(
  nlevels = seq(2, 10, 1),
  nsim = 500,
  create_fun = max
)
# sim_dist = distributional::dist_normal(5, 10))
# distr = "gumbel")
```

@@

```{r max_large_lev}
set.seed(54321)
plot_simdata(
  nlevels = seq(20, 45, 5),
  nsim = 500,
  create_fun = max
)
# distr = "gumbel",
# sim_dist = distributional::dist_normal(5, 10))
```

```{r norm_max}
# normalised maximum of a vector
norm_max <- function(x = NULL,
                     shape = shape1,
                      rate = rate1) {
  nx <- length(x)
  p <- 1 - (1 / nx)
  meanx <- shape / rate
  sdx <- shape / (rate^2)
  # meanx = 5
  # sdx = 10
  b <- stats::qnorm(p, meanx, sdx)
  a <- 1 / (nx * stats::dnorm(b, meanx, sdx))
  # k = stats::quantile(x, p, type = 8, na.rm = TRUE)
  (max(x) - a) / b
  # max(x)/k
}
```

@@

```{r normmax_small_lev}
set.seed(12345)
plot_simdata(
  nlevels = seq(2, 10, 1),
  nsim = 500,
  # distr = "gumbel",
  # sim_dist = distributional::dist_normal(5, 10),
  create_fun = norm_max
)
```

@@

```{r normmax_large_lev}
set.seed(54321)
plot_simdata(
  nlevels = seq(20, 45, 5),
  nsim = 500,
  create_fun = norm_max
  # distr = "gumbel",
  # sim_dist = distributional::dist_normal(5, 10))
)
```
